# Loading the packages
```{r libraries,results="hide"}
library(dplyr)
```

# Loading the data
```{r loading}
raw <- read.csv("activity.csv")
str(raw)
```

# Ignoring the missing data
```{r ignored.NAs}
ignored.NAs <- raw[!is.na(raw$steps),]
str(ignored.NAs)
```

# Total number of steps taken each day (ignoring the missing values)
Compute the total number of steps per day.
```{r total.per.day}
total.steps.per.day <- ignored.NAs %>%
                       group_by(date) %>%
                       summarize(steps = sum(steps))
str(total.steps.per.day)
```

Plot the total number of steps taken each day.
```{r histogram}
hist(total.steps.per.day$steps, main="Total number of steps taken each day", xlab="Total steps")
```

Compute the mean and median number of steps taken each day.
```{r mean.and.median}
mean1 <- mean(total.steps.per.day$steps)
print(mean1)
median1 <- median(total.steps.per.day$steps)
print(median1)
```

# Average number of steps taken (averaged across all days) versus the 5-minute intervals (ignoring the missing values)

Compute the mean numbers per interval.
```{r means.per.interval}
mean.steps.per.interval <- ignored.NAs %>%
                           group_by(interval) %>%
                           summarize(mean.steps = mean(steps))
str(mean.steps.per.interval)
```

Plot the mean numbers per internal.
```{r plot.means.per.interval}
with(mean.steps.per.interval, plot(interval, mean.steps, type='l'))
```

Compute the 5-minute interval that, on average, contains the maximum number of steps.
```{r max.steps.interval}
max.steps.value <- max(mean.steps.per.interval$mean.steps)
max.steps.interval <- mean.steps.per.interval$interval[mean.steps.per.interval$mean.steps==max.steps.value]
print(max.steps.interval)
```
The interval is `r max.steps.interval`.

# Imputing missing data

Get the count of missing steps values in raw data.
```{r check.missing.values}
missing.values <- sum(is.na(raw$steps))
print(missing.values)
```

Impute the mean numbers per interval (computed in a previous step) as the default values for missing values.
```{r impute}
imputed <- raw %>%
           left_join(mean.steps.per.interval) %>%
           mutate(steps = ifelse(is.na(steps), mean.steps, steps))

imputed$mean.steps <- NULL # Added by the left join, no more needed
```

# Total number of steps taken each day (after missing values were imputed)
Compute the total number of steps taken each day.
```{r total.per.day.imputed}
total.steps.per.day <- imputed %>%
                       group_by(date) %>%
                       summarize(steps = sum(steps))
str(total.steps.per.day)
```

Plot the total number of steps taken each day.
```{r histogram.imputed}
hist(total.steps.per.day$steps, main="Total number of steps taken each day", xlab="Total steps")
```

Compute the mean and median number of steps taken each day.
```{r mean.and.median.imputed}
mean2 <- mean(total.steps.per.day$steps)
print(mean2)
median2 <- median(total.steps.per.day$steps)
print(median2)
```

Compare with the mean and median numbers without imputing missing values:
```{r mean.and.median.compare}
print(mean1)
print(median1)
```
The mean remains the same (as we use the means for imputing, it looks consistent).

The median did change.

#  Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends (after missing values were imputed)

Add the daytype factor
```{r mutate.daytype}
# As I am not using the English locale, I prefer to use weekday numbers, rather than (localized) weekday names
day.of.week <- as.POSIXlt(imputed$date)$wday

weekend <- (day.of.week == 0) | (day.of.week == 6) # 0 = Sunday, 6 = Saturday

imputed.with.daytype <- imputed %>%
                        mutate(daytype = factor(weekend, labels=c("weekday", "weekend")))
```

Plot in 2 panels.

```{r plot.daytype}
# Prepare 2 panels
par(mfrow=c(2,1))

# Panel 1: weekend
mean.steps.per.interval <- imputed.with.daytype %>%
                           filter(daytype=="weekend") %>%
                           group_by(interval) %>%
                           summarize(mean.steps = mean(steps))

with(mean.steps.per.interval, plot(interval, mean.steps, type='l', main="weekend"))

# Panel 2: weekday
mean.steps.per.interval <- imputed.with.daytype %>%
                           filter(daytype=="weekday") %>%
                           group_by(interval) %>%
                           summarize(mean.steps = mean(steps))

with(mean.steps.per.interval, plot(interval, mean.steps, type='l', main="weekday"))
```
```{r plot.restore,echo=FALSE}
# Restore single panel
par(mfrow=c(1,1))
```

We have very different activity patterns. Much less steps in working days, at working hours.